---
title: "Are Movies Getting Worse?"
output:
  html_document:  
    keep_md: true
    toc: true
    toc_float: true
    code_folding: hide
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
---

```{r, include=FALSE}
library(mosaic)
library(DT) 
library(pander) #You'll love this one! 
# To use it Run this in your Console: install.packages("pander")
library(readr)
movies <- read_csv("../Data/movies.csv")
movies2 <- read_csv("../Data/movies2.csv")
# To get the movies and movies2 data into your Console, Run:
# "Session -> Set Working Directory -> To Source File Location"
# and then, "Play" this R-Chunk with the little green play button.
```

<br />

<!-- Instructions: Come up with an interesting question and answer it with the "movies" dataset that is provided above. 

Write a Background that gives meaning to your question of interest.

State your Question.

Perform an appropriate "Analysis" that gives useful information for answering your question. The hope would be a really powerful graphic and some summary tables of supporting statistics.

Interpret your analysis and expound on your findings. -->
#Background
In the world we live in today, it is thought the world is only getting worse. Modern-day ecclesiastical leaders have told us that the generation of today is a "Chosen Generation", because of the challenges they have today and the duties they have yet to fulfill. It's often difficult to really measure how much worse society is morally becoming. With data from every movie released since 1986, the folowing question was asked:'Have movies been progressing towards less family friendly productions?' Answering this question can reveal more about society and what is becoming more popular among the world.

#Analysis


```{r message=FALSE, warning=FALSE}
library(mosaic)
library(plotly) 
library(tidyr)
jose<-subset(movies2, rating=="G"|rating=="PG"|rating=="PG-13"|rating=="R")
juan<-prop.table(table(jose$rating, jose$year), 2)
joan<-data.frame(juan)
john<-spread(joan, Var1, Freq)
plot_ly(john, x=~Var2, y=~R, name = 'R', type = 'scatter', mode = 'lines', text=~paste(Var2)) %>% add_trace(y = ~`PG-13`, name = 'PG-13', mode = 'lines') %>%  add_trace(y = ~PG, name = 'PG', mode = 'lines') %>% add_trace(y = ~G, name = 'G', mode = 'lines')%>% layout(title="Comparisons of Movie Ratings by Proportion \n Since 1986", xaxis=list(title="Year"), yaxis=list(title='Yearly Proportion of Movies'))
```

I thought that the best way to decide whether or not movies were getting worse would be to find how the yearly proportion of movies created under each rating. The most important changes are in the statistics of PG, PG-13, and R movies. Note that these percentages are to compare the yearly amount of movies created in that rating compared to the other ratings in this study.The average percentage of movies that are PG-13 over the first 15 years is 24.8%, whereas the average percentage of movies that are PG-13 over the later 16 years is 36.5%. The average percentage of movies that are PG over the first 15 years is 17.3%. More recently, the average percentage of movies that are PG over the later 16 years is 12.1%. The average percentage of movies that are R over the first 15 years is 54.9%. More recently, the average percentage of movies that are R over the later 16 years is 49.9%. G rated movies don't have a significant change, though it should be noted it does decrease.

#Conclusion
For the case of this study, it's considered that PG and G movies are family-friendly, whereas PG-13 and R rated movies are not family-friendly. Despite the rated R movies percentage also decreasing slightly, it's visible that not family-friendly movies were increasing in comparison to family-friendly movies decreasing over the span of this time. This  supports the theory that family friendly movies are becoming less common. Additionally, it upholds the claims of the ecclesiastical leaders that these modern days will only be more challenging to choose the right.
  The overall conclusion is that moral people should be more cautious of the films they choose. One of the challenges faced in these modern times is that the increasing amounts of evil lower the standard for choosing the right. Non family-friendly movies are becoming much more common, however this shouldn't allow morally clean individuals to rationalize viewing them. All in all, lowering personal ethics to view more prevalent unclean material would be calling evil good; a popular mistake made to comfort justification.



[Details](https://vincentarelbundock.github.io/Rdatasets/doc/ggplot2/movies.html) for the movies dataset.





