---
title: "Increase in Heights Since 19th Century"
output:
  html_document:  
    theme: flatly
    code_folding: hide
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load_libraries, include=FALSE}
# Use this R-Chunk to load all your libraries!
#install.packages("tidyverse") # run this line once in console to get package
library(tidyverse)
library(haven)
library(tidyr)
library(foreign)
library(forcats)
library(downloader)
library(readr)
library(readxl)
library(ggthemes)
```

```{r load_data}
# Use this R-Chunk to import all your datasets!
wisc <- read_spss("http://www.ssc.wisc.edu/nsfh/wave3/NSFH3%20Apr%202005%20release/main05022005.sav")
wisc2 <- as.tibble(wisc)

german <- read_dta("https://byuistats.github.io/M335/data/heights/germanconscr.dta")
bavar <- read_dta("https://byuistats.github.io/M335/data/heights/germanprison.dta")

wageheight <- read_csv("https://github.com/hadley/r4ds/raw/master/data/heights.csv")


worlddata <- tempfile(fileext = '.xlsx')
download("https://byuistats.github.io/M335/data/heights/Height.xlsx", mode = 'wb', destfile = worlddata)
worlddata <- read_excel(worlddata, skip = 2)

south <- read.dbf("~/Desktop/B6090.DBF")

```



```{r tidy_data}
# Use this R-Chunk to clean & wrangle your data!
worlddata2 <- worlddata%>%
  gather('1800':'2011', key = 'year_decade', value = 'heightscm')%>%
  filter(!is.na(heightscm))%>%
  mutate(heightsin = heightscm*0.3937008)%>%
  separate(year_decade, into = c('Century','decade'),sep = 2, remove = FALSE)%>%
  separate(decade, into = c('Decade', 'Year'), sep = 1)
  
wisc3 <- wisc2%>%
  mutate(heightsin = RT216F*12+RT216I, year_decade = 1900 + DOBY, heightscm = heightsin/0.3937008, study = '1900 National Survey')%>%
  select(year_decade, heightsin, heightscm, study)

bavar2 <- bavar%>%
  mutate(year_decade = bdec, heightscm = height, heightsin = heightscm*0.3937008, study = '19th Century Bavarian')%>%
  select(year_decade, heightsin, heightscm, study)

german2 <- german%>%
  mutate(year_decade = bdec, heightscm = height, heightsin = height*0.3937008, study = '19th Century Bavarian Germans')%>%
  select(year_decade, heightsin, heightscm, study)

wageheight2 <- wageheight%>%
  mutate(year_decade = 1950, heightsin = height, heightscm = heightsin/0.3937008, study = '1950 BLS')%>%
  select(year_decade, heightsin, heightscm, study)

south2 <- south%>%
  mutate(year_decade = GEBJ, heightscm = CMETER, heightsin = CMETER*0.3937008, study = '19th Century South German Soldiers')%>%
  select(year_decade, heightsin, heightscm, study)

alld <- bind_rows(bavar2, wisc3, german2, wageheight2, south2)

alld$years <- cut(alld$year_decade, c(1700, 1800, 1900, 2000), labels = c('18th Century','19th Century','20th Century'))

```

## Data Visualization

```{r plot_data}
# Use this R-Chunk to plot & visualize your data!

world <- worlddata2%>%
  mutate(Country = case_when(
      `Continent, Region, Country` == 'Germany' ~ "Germany", 
      TRUE ~ 'Other'
      ))%>%
  group_by(year_decade, Country)%>%
  summarise(countrymean = mean(heightsin))
  
    
worlddata2%>%
    mutate(Country = case_when(
      `Continent, Region, Country` == 'Germany' ~ "Germany", 
      TRUE ~ 'Other'
      ))%>%
    ggplot()+
    geom_point(mapping = aes(x = year_decade, y = heightsin, color = Country))+
    geom_line(data = world, aes(x = year_decade, y = countrymean, group = Country, color = Country))+
  scale_color_manual(values = c('dodgerblue','darkgray'))+
  labs(title = "Germany's Increase in Height is More Extreme Than Rest of World", x = 'Year', y = 'Heights(in Inches)')+
  theme_stata()

alld%>%
  filter(heightsin<96 & heightsin>45&!is.na(years))%>%
  dplyr::mutate(study = fct_relevel(study, c('19th Century South German Soldiers','19th Century Bavarian Germans','19th Century Bavarian','1900 National Survey','1950 BLS')))%>%
  ggplot()+
  geom_jitter(mapping = aes(y = heightsin, x = study, color = study))+
  labs(title = 'No Large Increase in Heights since 18th Century', x = 'Century',y = 'Heights (in inches)')+
  theme_stata()+
  guides(fill = FALSE)
```


The first plot I created helps to describe the overall increase in heights over the years. Germany is highlighted, and according to the visualization, had a muc faster increase in heights over most other countries. The 'Other' variable in this plot shows all other countries or territories. 

My second plot is a visualization to compare all the other plots over the past few centuries. They are all in order of the years they were taken. Evidently, it appears as though there hasn't been much change. One could argue that there seem to be both taller and shorter people more recently, giving us more variance over the spectrum.