---
title: "Day Care Fines"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

<script type="text/javascript">
 function showhide(id) {
    var e = document.getElementById(id);
    e.style.display = (e.style.display == 'block') ? 'none' : 'block';
 }
</script>

```{r, include=FALSE}
library(mosaic)
library(DT)
library(pander)
library(car)

# To get the "late" dataset into your R Console, 
# you need to go to your file menu of R-Studio and select
# "Session, Set working directory, To source file location"
# Then click the green "play" arrow in 
# the top right corner of this gray R-Chunk. 
late <- read.csv("../Data/late.csv", header=TRUE)
# Then type:
# > View(late)
# in your R Console
```



----

<!--If you want to give your critiquers some ideas about what you have questions on and would like help with, place those details here.-->

**Comments to Critiquers:** 



<!--End your comments to your critiquers before this line.-->

----


## Background {.tabset .tabset-fade}

### Overview
<!-- Be sure to add your own background here before beginning the Analysis. -->

This analysis is based off of the study called 'A Fine is a Price', which studies the patterns of the amounts of kids that are picked up late from preschool. There is a period of 4 weeks where they record the amount of kids that are picked up late, followed by a period of 12 weeks where they fine those that pick up their kids late and record the amount, and finally a period of 4 weeks where they revoke the fine.


This study is important for finding out methods of helping day cares become more efficient. Teachers give valuable time when kids stay late and aren't paid for this. Thus, could fining families who pick their kids up late help establish a culture where they're picked up on time?


<!--End your background before here.-->
<!--Click the little black arrow to 
the left to expand/hide these next sections.-->
### Study Details

<div style="color:#a8a8a8;">
This background is quoted directly from the article ["A Fine is a Price"](http://rady.ucsd.edu/faculty/directory/gneezy/pub/docs/fine.pdf). 
</div>

<div style="padding-left:30px; padding-right:30px;">

There are two types of day-care centers in Israel: private and public. A study was conducted in 10 private day-care centers in the city of Haifa from January to June 1998. All of these centers are located in the same part of town, and there is no important difference among them. During the day children are organized into groups according to age, from 1 to 4 years old. Each day-care center is allowed to hold a maximum of 35 children. In some exceptional cases a few additional children are allowed. The fee for each child is NIS 1,400 per month. (The NIS is the New Israeli Shekel.) At the time of the study, a U.S. dollar was worth approximately NIS 3.68, so the fee was about $380 at that time. 

The contract signed at the beginning of the year states that the day-care
center operates between 0730 and 1600. There is no mention of what happens if parents come late to pick up their children. In particular, before the beginning of the study, there was no fine for coming late. When parents did not come on time, one of the teachers had to wait with the children concerned. Teachers would rotate in this task, which is considered part of the job of a teacher, a fact that is clearly explained when a teacher is hired. Parents rarely came after 1630. 

A natural option [to fix the problem of parents showing up late] is to introduce a fine: every time a parent comes late, [they] will have to pay a fine. Will that reduce the number of parents who come late? If the fine is removed, will things revert back to the way they were originally?

The overall period of the study was 20 weeks. In the first 4 weeks we
simply recorded the number of parents who arrived late each week. At the
beginning of the fifth week, we introduced a fine in six of the 10 day-care centers, which had been selected randomly. The announcement of the fine was made with a note posted on the bulletin board of the day-care center. Parents tend to look at this board every day, since important announcements are posted there. The announcement specified that the fine would be NIS 10 for a delay of 10 minutes or more. The fine was per child; thus, if parents had two children in the center and they came late, they had to pay NIS 20. Payment was made to the principal of the day-care center at the end of the month. Since monthly payments are made to the owner during the year, the fines were added to those amounts. The money was paid to the owner, rather then to the teacher who was staying late (and did not get any additional money). The teachers were informed of the fine but not of the study. Registering the names of parents who came late was a common practice in any case.

At the beginning of the seventeenth week, the fine was removed with no explanation. Notice of the cancellation was posted on the board. If parents asked why the fines were removed, the principals were instructed to reply that the fine had been a trial for a limited time and that the results of this trial were now being evaluated.

A comparison with other fines in Israel may give an idea of the size of
the penalty that was introduced. A fine of NIS 10 is relatively small but not insignificant. In comparison, the fine for illegal parking is NIS 75; the fine for driving through a red light is NIS 1,000 plus penalties; the fine for not collecting the droppings of a dog is NIS 360. For many of these violations, however, detection and enforcement are low or, as in the case of dog dirt, nonexistent in practice. A baby-sitter earns between NIS 15 and NIS 20 per hour. The average gross salary per month in Israel at the time of the study was NIS 5,595.

</div>


### The Data (Wide)


The `late` Day Care Center data is shown here in the "wide data format". 

```{r, out.width=300}
#Show the full width of the "Wide" version of the late data:
pander(late, split.tables = Inf)
```


### The Data (Long)

The `Late` Day Care Center data is shown here in the "long data format".

```{r}
# This code reshapes the data into "long" format called Late.
# To get the "Late" dataset into your R Console, 
# you need to click the green "play" arrow in 
# the top right corner of this gray R-Chunk. 
# Then type:
#   > View(Late)
# in your R Console
Late <- reshape(late,
                varying = paste("Week",1:20, sep=""), 
                v.names = "No.ofLateChildren",
                timevar = "Week", 
                times = 1:20, 
                idvar = "Center",
                new.row.names = 1:200,
                direction = "long")
pander(Late)
```


 
##  
<!-- You can begin writing again after here.-->



## Analysis
We decided to do a Two-Way Anova test to find if there is any difference in the effect of the fine on the school.
We have three questions posed from this:
Does the treatment plan have any affect on the amount of students picked up late over the time period?


Factors: Control-c, Fined-f
$$
  H_0: \mu_c = \mu_f  = \mu
$$

$$
  H_a: \mu_i \neq \mu \ \text{for at least one} \ i
$$
Does the period of time have an effect on the amount of kids being picked up late?

$$
  H_0: \mu_b = \mu_d = \mu_a = \mu
$$
$$
  H_a: \mu_i \neq \mu \ \text{for at least one} \ i
$$
Is there a relationship between the fine and the amount of kids being picked up in each period of time?
$$
  H_0: \text{There is no effect of treatment between the time periods.}
$$

$$
  H_a: \text{there is an effect of treatment between the time periods.}
$$

The following table represents the Anova test performed.
```{r message=FALSE, warning=FALSE}
library(plyr)

Late$section<-mapvalues(Late$Week, 1:20, c(rep('Before',4), rep('Fine', 12), rep('After', 4)))
litaov<- aov(No.ofLateChildren ~ Treatment+section+Treatment:section, data=Late)
pander(summary(litaov))
```

We created tests of normality shown with the following plots. The following also show that the data is approximately normal.It is also evident that the variance is approximately constant with no outstanding problem for our test.

```{r}
par(mfrow=c(1,2))
plot(litaov, which=1:2)
```



The following graphs were created to better express the effectiveness of each hypothesis test. The first compares the results between Control and Fine groups. The second represents the Tiem Periods and their effect on the results. Finally, the last paragraph compares the effect of the time periods and the treatment. 


```{r message=FALSE, warning=FALSE}

xyplot(No.ofLateChildren ~ Treatment, data=Late, type=c("p","a"), ylab = 'Number of Children Picked Up Late', main = "Effects of Treatment on Kids Being Picked Up Late")
```

```{r message=FALSE, warning=FALSE}
x1<-factor(Late$section, c('Before', 'Fine', 'After'))
xyplot(No.ofLateChildren ~ x1, data=Late, type=c("p","a"), xlab='Period', ylab = 'Number of Children Picked Up Late', main = 'Effect of the Time Periods on Kids Being Picked Up Late')
```

```{r message=FALSE, warning=FALSE}
x1<-factor(Late$section, c('Before', 'Fine', 'After'))
xyplot( No.ofLateChildren ~ x1, data=Late, groups = Treatment, type=c("p","a"), auto.key = list(corner=c(1,1)), xlab='Period', ylab = 'Number of Children Picked Up Late', main = 'Weekly Amount of Kids Picked Up in Day Cares')
```


##Interpretation
All three tests have sufficient evidence to reject the null hypothesis. We can affirm that the treatment, time period and relationship between the two all have an effect on the amount of kids that are late. It's worthy to note that though we probably expected the fine to decrease the amount of kids being picked up late, it actually makes it less effective. We would advise avoiding the fine, as it has proven to only increase the amount of kids that are picked up late. We could imagine that adding the fine allows parents to believe that leaving your kids late is acceptable. 
