---
title: "My Simple Logistic Regression"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---


```{r, include=FALSE}
library(mosaic)
library(ResourceSelection)
library(DT)
library(pander)
mlbwins <- read.csv("../Data/mlbwins.csv", header=TRUE)
```


## Background
Every year 30 teams in Major League Baseball compete to be able to win the World Series. There's only one champion, however after the regular season ten different teams of the 30 make it to the playoffs. Each team invest hundreds of millions of dollars to get there. The question posed is how many wins should a team be shooting for in order to make it to the playoffs? Depending on each amount of wins the teams have, what is their probability of making it to the playoffs?

I'll use the model below to explain my reasoning.
$$
  P(Y_i = 1|\, x_i) = \frac{e^{\beta_0 + \beta_1 x_i}}{1+e^{\beta_0 + \beta_1 x_i}} = \pi_i
$$

I'd like to define the meaning of $i$:

* $Y_i=1$ would mean that the MLB team will make the Playoffs.
* $Y_i=0$ would mean that the MLB team fails to reach the Playoffs.
* $X_i$ will represent the amount of wins each team has at the end of the regular season.

It'll be crucial to understand that the amount of wins is meaningless when $\beta_1$ is zero. Thus, I'll use the following hypotheses

$$
H_0: \beta_1=0\\

H_a: \beta_1 \neq 0 
$$

<br />

## Data

```{r}
datatable(mlbwins,options = list(lengthMenu=c(5,10,30,100)))
```


<br />

## Analysis

The plot below helps us understand the Logistic Regression model.

```{r}
plot(Playoffs>0~Wins, data= mlbwins, xlim=c(50, 110), pch=16, col='chocolate1', xlab='Amount of Regular Season Wins', ylab='Probability of Reaching the Playoffs', main ='Amount of Wins Needed to Reach  MLB Playoffs')
curve(exp(-65.8038+.7526*x)/(1+exp(-65.8038+.7526*x)), from=40, to=110, add=TRUE, col='navy', lwd=3)
```

I've summarized the coefficients and the test with the following table. The data shws significant results as the p-value is 9.535e-07. Thus we cand assume that $\beta_1$ is not equal to zero.
```{r message=FALSE, warning=FALSE}
mlbglm<-glm(Playoffs>0~Wins, data=mlbwins, family=binomial)
pander(summary(mlbglm))
```


<br />


Also, it's important to test the effectiveness of the Logistic Regression test on this data, so I performed a Hosmer-Lemeshow goodness of Fit test. 


I received the following results:
```{r message=FALSE, warning=FALSE}
pander(hoslem.test(mlbglm$y, mlbglm$fitted.values))
```

As seen in this test, the p-value is .7805. This will condlude that our null hypothesis that the logistic regression model is appropriate for this data is true. Therefore, our results are meaningful.

<br />

## Interpretation
From our research above, we find that $\beta_1$ is equal to about .7526. This means that with every one increase in $x_i$, that the probability of the MLB team making it to the playoffs increases by a factor of $e^0.7526$ or 1.21. NOte that the amount of mlb wins being 0 in a season is not really relevant, so $\beta_0$ is not interpretable. Our resultant model is the following:
$$
  P(Y_i = 1|x_i) \approx \frac{e^{-65.8+.7526\cdot 31}}{1+e^{-65.8+.7526 \cdot x_1}} = \hat{\pi}_i
$$
In order for a team to have a 50% chance of making the playoffs, they will need to have about 87.5 Wins in their regular season.