---
title: "Income and Education on Political Party"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

```{r setup, include=FALSE}
GSS <- read.table("../Data/GSS2012.csv", sep="\t", header=TRUE)
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(mosaic)
library(car)
library(plyr)
library(DT)
library(pander)
GSS2012<-GSS
GSS2012$neweduc<-mapvalues(GSS2012$educ, c(0,1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20), c(rep('1', 12), '2', rep('3', 3), '4', rep('5', 4)))
GSSC<-GSS2012[c('coninc', 'neweduc', 'partyid', 'age')]
GSSC$newparty<-mapvalues(GSSC$partyid, c(0,1,2,4,5,6), c(rep('Democrat', 3), c(rep('Republican', 3))))
GSS2<-GSSC[c('coninc', 'neweduc', 'newparty', 'age')]
GSS3<-subset(GSS2, newparty=='Democrat'|newparty=='Republican')
GSS4<-subset(GSS3, neweduc!=98)
x1<-factor(GSS4$neweduc, c('Some HS or Less', 'HS Graduate', 'Some College', 'College Graduate',"Master's or Doctorate's Degree"))
litglm<-glm(newparty=='Democrat'~coninc+neweduc, data=GSS4, family=binomial)
b<-coefficients(litglm)
```
## Background
This analysis will be based off of the General Social Survey, which is a survey done for scientific research based on Americans and their development. I decided to test how the probability of being a democrat might change based off of income and education.


I will be using the following logistic regression equation to display my test:
$$
  P(Y_i = 1|\, x_{i1},x_{i22},\ldots,x_{i26}) = \frac{e^{\beta_0 + \beta_1 x_{i1} + \beta_2 x_{i22} + \ldots + \beta_6 x_{i26}}}{1+e^{{\beta_0 + \beta_1 x_{i1} + \beta_2 x_{i22} + \ldots + \beta_6 x_{i26}}}} = \pi_i
$$

For this analysis, $Y_i=0$ denotes the political party being Republican of that individual and $Y_i=1$ would denote that the political party is Democrat. To explain each different variable, I've created the following table:

| Variable | Value | Explanation |
|----------|-------|-------------|
| $x_{i1}$ | `0 to 180386` | the yearly income of the family |
| $x_{i22}$ | `neweduc == 1` | individual never finished High School |
| $x_{i23}$ | `neweduc == 2` | individual has graduated High School |
| $x_{i24}$ | `neweduc == 3` | individual has less than 4 years of college after High School |
| $x_{i25}$ | `neweduc == 4` | individual has 4 years of college after High School |
| $x_{i26}$ | `neweduc == 5` | individual has 5-8 years of college after High School |




My hypothesis to be tested by the logistic regression will be the following:
$$
  H_0: \beta_0 = 0 \\
  H_a: \beta_0 \neq 0
$$
$$
  H_0: \beta_2 = 0 \\
  H_a: \beta_2 \neq 0
$$
$$
  H_0: \beta_3 = 0 \\
  H_a: \beta_3 \neq 0
$$
$$
  H_0: \beta_4 = 0 \\
  H_a: \beta_4 \neq 0
$$
$$
  H_0: \beta_5 = 0 \\
  H_a: \beta_5 \neq 0
$$

$$
  H_0: \beta_6 = 0 \\
  H_a: \beta_6 \neq 0
$$

## Data

```{r message=FALSE, warning=FALSE}
datatable(GSS4,options = list(lengthMenu=c(5,10,30,100)))
```


## Analysis


I created the following plot to portray the variables and how they appear in relation to each other.

```{r message=FALSE, warning=FALSE}
plot(newparty=='Democrat'~coninc, data=GSS4, xlab='Family Income (in USD)', ylab='Probablility of Being a Democrat')
curve(exp(b[1]+b[2]*x)/(1+exp(b[1]+b[2]*x)), add=TRUE, col='cadetblue', lwd=2)
curve(exp(b[1]+b[3]+b[2]*x)/(1+exp(b[1]+b[3]+b[2]*x)), add=TRUE, col='chartreuse4', lwd=2)
curve(exp(b[1]+b[4]+b[2]*x)/(1+exp(b[1]+b[4]+b[2]*x)), add=TRUE, col='chocolate3', lwd=2)
curve(exp(b[1]+b[5]+b[2]*x)/(1+exp(b[1]+b[5]+b[2]*x)), add=TRUE, col='brown3', lwd=2)
curve(exp(b[1]+b[6]+b[2]*x)/(1+exp(b[1]+b[6]+b[2]*x)), add=TRUE, col='darkgoldenrod3', lwd=2)
legend("bottomright",legend=paste("neweduc ==", 1:5), col=c("cadetblue","chartreuse4","chocolate3","brown3","darkgoldenrod3"), lty=1)
```

This plot is based off the following table which presents the statistics from a Logistic Regression test.

```{r message=FALSE, warning=FALSE}
GSS2012<-GSS
GSS2012$neweduc<-mapvalues(GSS2012$educ, c(0,1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20), c(rep('1', 12), '2', rep('3', 3), '4', rep("5", 4)))
GSSC<-GSS2012[c('coninc', 'neweduc', 'partyid', 'age')]
GSSC$newparty<-mapvalues(GSSC$partyid, c(0,1,2,4,5,6), c(rep('Democrat', 3), c(rep('Republican', 3))))
GSS2<-GSSC[c('coninc', 'neweduc', 'newparty', 'age')]
GSS3<-subset(GSS2, newparty=='Democrat'|newparty=='Republican')
GSS4<-subset(GSS3, neweduc!=98)
litglm<-glm(newparty=='Democrat'~coninc+neweduc, data=GSS4, family=binomial)
b<-coefficients(litglm)
pander(summary(litglm))
```


</div>

In order to find if this logistic regression was relevant and appropriate, I performed a goodness of fit test. The null hypothesis is that logistic regression is appropriate. The p-value for the goodness of fit test is the following:

```{r message=FALSE, warning=FALSE}
pander(pchisq(2011.6, 1526, lower.tail = FALSE))
```

</div>

Thus, we see that there is sufficient evidence to reject the null hypothesis. A logistic regression is not apprpriate for this data. However, I will continue to do the interpretation to interpret what the data says if it was apprpriate.

## Interpretation
When we tested the different variables we found that only a few were significant. The test for College Graduates, those with Master's Degrees and Doctorate's Degrees, and those with some HS or less all had significant p-values; or in other words when neweduc = 1, 4 and 5. The income of the family is important to note just to know that as income increased by 1, the probability of an individual being Democrat decreased by $e^{-4.45e-06}=0.9999956$. 

Now for neweduc==3 or Master's or Doctorate's Degree, the variable is the effect that having a Master's or Doctorate's degree will have on the odds of being a democrat. The variable was $e^{0.8774}=2.40464$ or about 140% higher odds to be a democrat. 

The rest of the results can be interpretted similarly. The overall result would be that there is sufficient evidence to conclude that those with Master's or Doctorate's degrees tend to lean more towards being a Democrat over category of educational achievement.
