---
title: "Recalling Recalling Words"
output: html_document
---

```{r setup, include=FALSE}
library(mosaic)
library(car)
library(DT)
library(pander)
library(plotly)
datatable(Friendly)
```

## Background

In the `RecallingWords` analysis of Week 4 (Wilcoxon Tests) you were restricted to testing only two groups at a time. Revisit that analysis with a Kruskal-Wallis Test. What advantages and disadvantages does this approach have over your original analysis?

$$
H_0: \text{The samples are all from the same distribution}
$$
$$
H_a: \text{At least one sample's distribution is stochastically different}
$$
## Graphic

```{r}
sfr<-subset(Friendly, condition == 'SFR')
before<-subset(Friendly, condition == 'Before')
mesh<-subset(Friendly, condition == 'Meshed')
plot_ly(type='box') %>%
add_boxplot(y=sfr$correct, name = 'SFR') %>%
add_boxplot(y=before$correct, name = 'Before') %>%
add_boxplot(y=mesh$correct, name = 'Meshed') %>%
layout(title='Scores of Words Exam by Group', yaxis=list(title='Number of Words Recalled'), xaxis = list(title = 'Group'))
```


## Summary Table

```{r}
pander(favstats(sfr$correct))
pander(favstats(mesh$correct))
pander(favstats(bef$correct))
```


## Kruskal-Wallis Test

```{r}
pander(kruskal.test(correct~condition, data=Friendly))
```



## Interpretation



## Conclusion

Be sure to compare how the conclusions of this approach differ from your conclusions of your Wilcoxon Test analysis. Which approach do you find more meaningful, and why?